---
title: "nClimGrid"
author: "C Schultz"
date: "1/21/2020"
output: html_document
---

```{r setup, include=FALSE}
# -----------------------------------------------------------------------------
# -----------------------------------------------------------------------------
# Initial setup; key inputs
# -----------------------------------------------------------------------------
# -----------------------------------------------------------------------------
# Data processing and analysis basics
library(tidyverse)
library(dplyr)
library(RcppRoll)
library(data.table)
library(zoo)
library(ggplot2)
library(trend)
# For geospatial analysis
library(ncdf4)
# library(rhdf5)
# library(RCurl) # needed if password protected site

# GIS packages
library(maptools)
library(raster)
library(rgeos)
#
# source("code/functions/read_br_spatialave_func.R", local = TRUE)

knitr::opts_chunk$set(echo = TRUE)
```
# Intro

The dataset, nClimGrid, is cited in NCA4. This dataset, the Gridded 5km GHCN-Daily Temperature and Precipitation Dataset, Version 1 (GHCN is NOAA’s Global Historical Climatology Network). The data is from the NOAA National Centers for Environmental Information (NCEI), and is described on the webpage, https://data.nodc.noaa.gov/cgi-bin/iso?id=gov.noaa.ncdc:C00332. 

The citation is: Vose, Russell S., Applequist, Scott, Squires, Mike, Durre, Imke, Menne, Matthew J., Williams, Claude N. Jr., Fenimore, Chris, Gleason, Karin, and Arndt, Derek (2014): Gridded 5km GHCN-Daily Temperature and Precipitation Dataset (nCLIMGRID), Version 1. [indicate subset used]. NOAA National Centers for Environmental Information. DOI:10.7289/V5SX6B56 [access date].

The ftp site for the nClimGrid data is ftp://ftp.ncdc.noaa.gov/pub/data/climgrid/. The data extends from 1895 to 2013 (then files with “preliminary” in the name up to current day). This is monthly data of total precipitation (mm), and mean, min and max temperature (deg C) for the CONUS at 5 km spatial resolution. 

## Don't want to lose these references:

http://geog.uoregon.edu/bartlein/courses/geog490/week04-netCDF.html

https://www.earthdatascience.org/courses/earth-analytics/time-series-data/

```{r, echo=FALSE}
# -----------------------------------------------------------------------------
# -----------------------------------------------------------------------------
# Preliminaries
# -----------------------------------------------------------------------------
# -----------------------------------------------------------------------------

# Read in Potomac watershed ---------------------------------------------------
upper_pot_shp_file <- "input/gis/UpperPotWatershed_dd.shp"
upper_pot_poly <- readShapePoly(upper_pot_shp_file)
# summary(upper_pot_shp)
# x -79.69099 -76.89138
# y  37.87822  40.23735
upper_pot_poly <- as(upper_pot_poly, "SpatialPolygons")

# Download 1st nClimGrid file -------------------------------------------------
nclimurl <- "ftp://ftp.ncdc.noaa.gov/pub/data/climgrid/"
filename <- "nClimGrid_v1.0_monthly_189501_c20150401.tar.gz"
nclimfile <- paste(nclimurl, filename, sep="")
destfile <- paste("data/nclimgrid_5km_v1.0/", filename, sep="")
download.file(nclimfile, destfile, method = "libcurl", quiet = FALSE, mode = "wb")

# Unzip the file --------------------------------------------------------------
untar(destfile, exdir = "data/nclimgrid_5km_v1.0/nclim189501/")

# Read file and clip to rectangle ---------------------------------------------
tmppath <- "data/nclimgrid_5km_v1.0/nclim189501/"
tmpname <- "189501.prcp.conus.pnt"
tmp_precip.df <- file.path(paste(tmppath, tmpname, sep="")) %>%
    data.table::fread(
      data.table = FALSE,
      header = FALSE,
      showProgress = FALSE, col.names = c("lat", "lon", "p_mm")) %>%
  filter(lat > 37.8 & lat < 40.3 & lon > -79.8 & lon < -76.8)

# Create nClimGrid coordinates and pnts objects -------------------------------
nclimgrid_coords <- cbind(tmp_precip.df$lon, tmp_precip.df$lat)
nclimgrid_pnts <- SpatialPoints(coords = nclimgrid_coords)
upper_pot_5km_pnts <- over(nclimgrid_pnts, upper_pot_poly)
upper_pot_5km_pnts2 <- data.frame(nclimgrid_coords[!is.na(upper_pot_5km_pnts), ])
names(upper_pot_5km_pnts2) <- c("lon", "lat")

# Create empty df for upper Pot monthly means ---------------------------------
upper_pot_monthly <- data.frame(year = integer(), month = integer(),
                                prcp_mm = numeric(), tave_degC = numeric())

# Delete *.tar.gz file --------------------------------------------------------
unlink(paste("data/nclimgrid_5km_v1.0/", filename, sep=""))
unlink("data/nclimgrid_5km_v1.0/nclim189501", recursive = TRUE)

# plot(upper_pot_poly)
# points(upper_pot_5km_pnts2)

```

```{r}
# -----------------------------------------------------------------------------
# -----------------------------------------------------------------------------
# Download and process all files
# -----------------------------------------------------------------------------
# -----------------------------------------------------------------------------
for(iyear in 1895:1930) {
  for(imonth in 1:12) {
    cmonth <- c("01", "02", "03", "04", "05", "06", "07", "08", "09",
                "10", "11", "12")
    # Download nClimGrid file -------------------------------------------------
    nclimurl <- "ftp://ftp.ncdc.noaa.gov/pub/data/climgrid/"
    filename <- paste("nClimGrid_v1.0_monthly_", iyear, cmonth[imonth],
                      "_c20150401.tar.gz", sep="")
    nclimfile <- paste(nclimurl, filename, sep="")
    destfile <- paste("data/nclimgrid_5km_v1.0/", filename, sep="")
    download.file(nclimfile, destfile, method = "libcurl", 
                  quiet = FALSE, mode = "wb")
    
    # Unzip the file ----------------------------------------------------------
    newdir <- paste("data/nclimgrid_5km_v1.0/nclim", iyear, 
                    cmonth[imonth], "/", sep="")
    untar(destfile, exdir = newdir)
    
    # Read files ---------------------------------------------
    # tmppath <- "data/nclimgrid_5km_v1.0/nclim189501/"

    # prcp_fname <- "189501.prcp.conus.pnt"
    prcp_fname <- paste(iyear, cmonth[imonth], 
                        ".prcp.conus.pnt", sep="")
    # tave_fname <- "189501.tave.conus.pnt"
    tave_fname <- paste(iyear, cmonth[imonth], 
                        ".tave.conus.pnt", sep="")
    
    tmp_prcp.df <- file.path(paste(newdir, prcp_fname, sep="")) %>%
      data.table::fread(data.table = FALSE, header = FALSE,
                        showProgress = FALSE, 
                        col.names = c("lat", "lon", "p_mm")) %>%
      filter(lat > 37.8 & lat < 40.3 & lon > -79.8 & lon < -76.8)
    
    tmp_tave.df <- file.path(paste(newdir, tave_fname, sep="")) %>%
    data.table::fread(
      data.table = FALSE,
      header = FALSE,
      showProgress = FALSE, col.names = c("lat", "lon", "tave_degC")) %>%
      filter(lat > 37.8 & lat < 40.3 & lon > -79.8 & lon < -76.8)
    
    # Clip file data to upper Pot ---------------------------------------------
    fin_prcp.df <- left_join(upper_pot_5km_pnts2, tmp_prcp.df, 
                             by = c("lon", "lat"))
    fin_tave.df <- left_join(upper_pot_5km_pnts2, tmp_tave.df, 
                             by = c("lon", "lat"))
    
    # Compute upper Pot means -------------------------------------------------
    upper_pot_prcp <- summarise(fin_prcp.df, mean(p_mm))
    upper_pot_tave <- summarise(fin_tave.df, mean(tave_degC))
    
    # Add to upper_pot_monthly
    newrow <- data.frame(year = iyear, month = imonth, 
                         prcp_mm = fin_prcp.df[1, 3],
                         tave_degC = fin_tave.df[1, 3])
    upper_pot_monthly <- rbind(upper_pot_monthly, newrow)
    
    # Cleanup
    unlink(paste("data/nclimgrid_5km_v1.0/", filename, sep=""))
    newdir2 <- paste("data/nclimgrid_5km_v1.0/nclim", iyear, 
                    cmonth[imonth], sep="")
    unlink(newdir2, recursive = TRUE)
  }
}

write_csv(upper_pot_monthly, "output/upper_pot_monthly.csv")
```

## 

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
